---
title: "Caravans - Modelling"
output: html_notebook
---

## 0. Setup

Load packages.

```{r packages, warning = FALSE, message= FALSE}
library(tidyverse)
library(tidymodels)
library(caret)
library(ISLR)
library(GGally)
library(plotly)
```

## Caravan data

Save Caravan data as a tibble.

```{r caravan_df}
caravan_df <- as_tibble(Caravan)

glimpse(caravan_df)

# Number of NA values
caravan_df %>%
  select(-Purchase) %>%
  summarise_all(~sum(is.na(.))) 

# Minimum values
caravan_df %>%
  select(-Purchase) %>%
  summarise_all(min) 

# Maximum vaues
caravan_df %>%
  select(-Purchase) %>%
  summarise_all(max) 

# Unique values for each variable
caravan_df %>%
  map(~sort(unique(.)))

```

Read in subset of variables for modelling from EDA analysis.
```{r var_mod}
var_mod <- read_csv("var_mod.csv")

var_mod <- var_mod$x
```

## Split into training and test datasets

Split 70% of the data into a training set and the other 30% as a test set.

```{r train_test}
set.seed(12)
split <- caravan_df %>%
  # select(one_of(var_mod), Purchase) %>%
  initial_split(prop = 0.7)

caravan_train <- training(split)
caravan_test <- testing(split)

# Specify recipe for pre-processing data and prep train and test datasets using this recipe.
rec <-caravan_train %>%
  recipe( ~ .) 

prepped <- rec %>%
  prep(retain = TRUE)
  
train <- prepped %>% 
  juice()

test <- prepped %>% 
  bake(new_data = caravan_test)
```

## Purchase classification model using boosting

Start by using the entire dataset, and see if the variables which appear to be the most important are consistent with those identified in the EDA.

```{r entire_boost}
# Specify the process to select the hyperparameters
fitControl <- trainControl(method = "repeatedcv",
                           number = 10, repeats = 5, classProbs = TRUE)

# Fit boosting model using 'train'
boost_mod <-
  train(
    Purchase ~ . ,
    data = train,
    method = "gbm",
    trControl = fitControl,
    verbose = FALSE
  )

boost_mod

# Give prediction probabilities on test set
test <- test %>%
  bind_cols(predict(boost_mod, newdata = test, type = "prob")) %>%
  mutate(pred_purchase = if_else(Yes >= 0.2, "Yes", "No"),
         pred_purchase = factor(pred_purchase))
```


```{r entire_boost_performance}
# a. Tree model accuracy compared to null model. Accuracy is % of predicted classifications that are the same as the observed.
accuracy(test, truth = "Purchase", estimate = "pred_purchase")

# b. Confusion matrix
cs_tree_conf_matrix <- test %>% 
  yardstick::conf_mat(truth = "Purchase", estimate = "pred_purchase") %>%
  .$table

cs_tree_conf_matrix %>%
  as_tibble() %>%
  ggplot(aes(Prediction, Truth, alpha = n)) + 
  geom_tile(fill = "darkblue", show.legend = FALSE) +
  geom_text(aes(label = n), color = "black", alpha = 1, size = 8) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(
    title = "Confusion matrix"
  )
```











